---
title: "🗓️ Session 10: Linear regression and experiments in R"
date: "2025-12-05"
date-modified: last-modified
---

This session focuses on the practical implementation of linear regression models
and on how to conduct the most common tasks associated with
designing and implementing experiments in R.

This way it complements the more theoretical lectures. More precisely, we focus
on the following topics:

- Implement simple and multiple linear regression in R
- Calculate and interpret $R^2$ manually and using `R` functions
- Understand and demonstrate omitted variable bias
- Perform data transformations to linearize relationships
- Create effective visualizations using `geom_smooth()`
- Conduct basic model diagnostics
- Implement a power analysis in R

## 📚 Mandatory Reading

- The tutorial [Linear Regression in R](/content/tutorials/linear-regression/index.qmd) 
  - [Data sets used](/content/tutorials/linear-regression/RegressionDataSets.zip)
- The tutorial [Experiments in R](/content/tutorials/experiments/index.qmd)
  - [Data sets used](/content/tutorials/experiments/ExperimentsDataSets.zip)

## Further Reading

- [Chapter 5](https://moderndive.com/v2/regression.html) and [Chapter 6](https://moderndive.com/v2/multiple-regression.html) in @moderndive

## In-class exercises

- [Work sheet](content/material/session10/student_template)^[If it does not open to download, please use right-click and then "Save as"]
- [Data sets](content/material/session10/session10_data.zip)

## ✍️ Coursework

- Complete the in-class exercises at home 
- Do the exercises `LinearRegression1` and `LinearRegression2` from the `DataScienceExercises` package

::: {.callout-tip collapse="true"}
## Quick code for starting the exercises I
```{r}
#| eval: false
learnr::run_tutorial(
  name = "LinearRegression1", 
  package = "DataScienceExercises", 
  shiny_args=list("launch.browser"=TRUE))
```
:::

::: {.callout-tip collapse="true"}
## Quick code for starting the exercises II
```{r}
#| eval: false
learnr::run_tutorial(
  name = "LinearRegression2", 
  package = "DataScienceExercises", 
  shiny_args=list("launch.browser"=TRUE))
```
:::

